SHIV: Reducing Supervisor Burden using Support Vectors to Efficiently Learn Robot Control Policies from Demonstrations [v17]

### 2015-09-01 paper.tex
added Florian's comments

\title{SHIV: Reducing Supervisor Burden using Support Vectors to Efficiently Learn Robot Control Policies from Demonstrations [v16]}

\author{Michael Laskey$^1$, Sam Staszak$^3$, Wesley Yu-Shu Hsieh$^1$, Jeff Mahler$^1$, Florian T. Pokorny$^1$, Anca Dragan$^1$, Ken Goldberg$^{1-2}$% <-this % stops a space
\thanks{$^1$ Department of Electrical Engineering and Computer Sciences; {\small \{mdlaskey,iamwesleyhsieh, ftpokorny, anca\}@berkeley.edu}}%
\thanks{$^2$ Department of Industrial Engineering and Operations Research; {\small goldberg@berkeley.edu}}%
\thanks{$^{1-2}$ University of California, Berkeley;  Berkeley, CA 94720, USA}%
}
\begin{document}

@@

In model-free robot learning from demonstration, a robot learns to perform a task, such as driving or grasping an object in a cluttered environment as shown in Fig. \ref{fig:teaser}, from examples provided by a competent supervisor, usually a human. The problem is model-free if  the system dynamics are unknown, eg. dynamics model may be highly non-linear or noisy. In learning from demonstration, the cost function, which is the negative of the reward function, is usually unknown or significantly delayed. In the grasping for clutter example, shown in Fig. \ref{fig:teaser}(b) the robot is only penalized when an object is knocked off the table, which means a potentially large number of states can be visited before any feedback. To avoid this a supervisor provides guidance through demonstrations and the robot learns a policy mapping states to controls by observing~\cite{argall2009survey}.

@@

In SHIV, we use a query selection method that allows for a non-stationary distribution, evaluating risk based on \emph{novelty} (is the state an outlier to the training distribution), as in Fig. \ref{fig:support_example}(c). Additionally, we also assign higher risk to  historically mislabeled regions, as in Fig. \ref{fig:support_example}(d). We then use a classification method based on the One Class SVM \cite{scholkopf2001estimating} to predict if a state is risky or not risky, which we we define as safe.%Our approach builds on the One Class SVM method for approximating quantile level sets \cite{}, which has already been used in novelty detection on high-dimensional image data \cite{}.

@@

We assume a known state space and set of controls. We assume access to a robot or simulator, such that we can start in some initial state and apply a series of controls and observe the following states that occur.  Lastly, we assume access to a supervisor that can provide a series of demonstrated trajectories initially and can be given a state visited by the robot or simulator and provide a control signal of what the supervisor would do.

@@

Providing correct control inputs or ``labeling'' at each
iteration and each state in DAgger can be tedious, imposing a large burden on the supervisor.
Instead of asking the supervisor for labels at all visited states $\{\bx_0,..,\bx_T\}$, SHIV uses a measure of \emph{risk} to actively \emph{decide} whether a label is necessary. For every state encountered $\bx_t$, it applies a \emph{decision function} $g_{\sigma}$, which we introduce in the next section, parametrized by a risk-sensitivity parameter $\sigma$ which will capture how smooth we expect the policy function to be: smoother functions imply that training examples have larger support, and data nearby will be less risky.

SHIV only asks for supervision on states for which $g_{\sigma}(\bx)=-1$:

@@

\vspace{-2ex}
\begin{align}\label{eq:decision_func}
g_{\sigma}(\bx) = \mbox{sgn}(\langle w,\Phi(x) \rangle-\rho).
\end{align}

@@

where $\alpha_i$ corresponds to the dual variables.    For training data in $\mathcal{X}$, we consider the quantile super-levelset for a user-defined threshold $\mu$ as a natural region within which
a trained policy $\pi_{\theta}$ can likely be learned since a sufficient data-density is available.
However, even when sufficient data is available, the associated control inputs may be inconsistent or noisy and a resulting policy
optimizing Eq. \ref{eq:super_objj} may still incur a large surrogate loss. To account for this, we propose a
modification to the One Class SVM:

@@

We then divide up our data in two sets those correctly classified:
$$D_{g}=\{\lbrace \bx_i,\bu_i \rbrace \in \mathcal{D}_{k}, y=1\}$$
and those states incorrectly classified: 
$$D_{b}=\{\lbrace \bx_i,\bu_i \rbrace \in \mathcal{D}_{k}, y=-1\}$$
A separate One-Class SVM is then trained on each set of states, ($D_{g}$ and $D_{b}$) and providing measures of the level sets, $g_{\sigma_g}$ and $g_{\sigma_b}$.

We then define the overall decision function as:
$$g_{\sigma}(\bx) = g_{\sigma_g} (\bx)+ g_{\sigma_b}(\bx)$$
points are deemed risky if $g_{\sigma}(\bx) \neq 0$.  Practically, this modification corresponds to
`carving out holes' in the estimated quantile super-levelset such that neighborhoods around states with $y_i=-1$ are
excluded from the super-levelset.

@@

Our modified One Class SVM contains two different $\nu$ parameters, $\nu$ and $\nu_g$. We set $\nu = 1e-3$ and $\nu_g = 1e-1$ for all experiments. The intuition for this is that 'hole-punching' One Class SVM, $g_{\sigma_b}$, should treat all points as inliers, to be conservative. The level set estimation should be robust to outliers in the dataset thus only looks at $90\%$ of points. We tuned $\sigma$ and $\sigma_b$ by doing a grid search over different values on the surrogate loss for a single trial of SHIV for 3 iterations for each example. \mlnote{we start the grid search with a rule of thumb that it should scale with feature size, hence the car example is 200 while the grasping is 5 and the needle is 2, looking for a citation on this though. Scikit learn does the same thing for kenerlized regression}

@@

The supervisor drives around the track twice. We collect raw images of the simulation from a 2D bird's eye view
and use Gaussian Pyramids to down-sample the images to $125 \times 125$ RGB pixels and then extract Histogram of
Oriented Gradients (HOG) features using OpenCV. This results in a $27926$ dimensional state space description.
For both DAgger and \acro, we use a Linear Support Vector Machine (SVM) to parameterize allowable
policies $\pi_{\theta}$, with $\gamma=0.01$ as a regularization term on the slack variables, which was set via cross
validation on the initial training examples. For the optimization of Eq. \ref{eq:dual_sup} in \acro, we used an
exponential kernel with a bandwidth of $\sigma_g = 200$ and $\sigma_b = 200$ to account for $95\%$ of the training distribution.

@@

\textbf{Comparison to active learning approaches:}
To compare our method against a traditional confidence based active learning approach, we use a confidence measure  of distance from the hyperplane \cite{tong2002support}. The threshold was set to the average distance from the hyperplane for the mis-classified points in $\mathcal{D}_0$, which consisted of two demonstrations from our solver. We further compared against Query By Committee, which was trained via bagging \cite{breiman1996bagging}. Our data set was uniformly divided up into 3 overlapping subsets each with $80\%$ of the data then for each subset a Linear SVM was trained on. If the three classifiers agreed with each other the point was determined low risk and if they disagree it was determined high risk. We finally compared our combined risk approach against the traditional One Class SVM without the "hole-punching". We measured the performance in terms of how many states are estimated to be of low risk (i.e. correctly classified) and are of actually high risk on the first policy roll out. Results  are averaged over 50 tracks and shown in Fig. \ref{fig:active_comp}.

@@

\begin{figure}[t]
\centering
\includegraphics[width=9cm, height=4cm]{figures/risk_bar.eps}
\caption {  A comparison of different active learning approaches in correctly classifying states to be of high risk during the initial policy roll out in the driving experiment . We compare against a confidence estimate of distance to hyperplane, query by committee for 3 hypothesis classifiers, the One Class SVM, and our modified One Class SVM.  Results  are averaged over 50 tracks and the policy $\pi_\theta$ is represented as a Linear SVM. 
   }

@@

\begin{figure}[t!]
\centering
\includegraphics[width=8cm, height=6cm]{figures/dagger_shiv_one_class.eps}
\caption{We compare performance in terms of minimization of the underlying cost function $c(\bx,\bu)$, which is the  number of times the car left the track versus the number of queries made to the supervisor. In Fig. \ref{fig:car_cost}, we plot the performance of DAgger and SHIV.  Initial results, which are run for 6 iterations each and are averaged over 40 levels, shown in Fig. \ref{fig:car_cost} suggest an $70\%$ reduction in the number of queries needed for SHIV compared to DAgger,}
\vspace*{-10pt}
\label{fig:car_cost}
\end{figure}

@@

A state $\bx\in \mathcal{X}$ consisted of the 3 dimensional pose of the six objects on the table (translation and rotation), the 3 joint angles of the arm and a scalar value in the range $[0,1]$ that measured the position of the gripper, $1$ being fully closed and $0$ being opened. For our representation of $\pi_{\theta}$, we used kernelized ridge regression with the radial basis function as the kernel with the default Sci-Kit learn parameters. We defined the cost functions, $C(\bx,\bu)$, as the sum of the number of objects knocked off the table plus $10$ times the binary value indicating if the object is grasped or not. The order of magnitude difference in cost for grasping the object is to place emphasize on that portion of the task. The bandwidth parameters for SHIV were set to $\sigma_b = 5$ and $\sigma = 6$. For the $\epsilon$ term in the our risk method, we used the median in regression error which was the L2 distance between the predicted control and the true supervisor's control.

@@

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth, height=6cm]{figures/grasp_clutter.eps}
\caption{We compare performance in terms of minimization of the underlying cost function $c(\bx,\bu)$, which is euclidean distance between translation in centimeters. In Fig. \ref{fig:car_cost}, we plot the performance of DAgger and SHIV without the modification.  Initial results, which are run for 20 iterations each and are averaged over 40 different initial starting positions, shown in Fig. \ref{fig:car_cost} suggest an $67\%$ reduction in the number of queries needed for SHIV compared to DAgger,}
\vspace*{-10pt}
\label{fig:grasp_cost}
\end{figure}

@@

For trials we sample a start position from a Gaussian on the translational component of $P_0$ with isotropic variance of
$0.1$ cm. The distance between initial $P_0$ and $P_G$ is roughly $3$ cm. Our cost function $c(\bx,\bu)$ is measured in
terms of Euclidean distance in translational component, which is in centimeters. \mlnote{We can do Frobenius norm on transformation matrix, wanted to double check if that was accurate though} We run DAgger and SHIV both for 20 iterations and average over 40 different starting positions.
Results shown in Fig. \ref{fig:needle} suggest a reduction $67\%$ in the amount of human inquiries needed.

>>>>

\title{SHIV: Reducing Supervisor Burden using Support Vectors to Efficiently Learn Robot Control Policies from Demonstrations [v17]}

\author{Michael Laskey$^1$, Sam Staszak$^3$, Wesley Yu-Shu Hsieh$^1$, Jeff Mahler$^1$, Florian T. Pokorny$^1$, Anca Dragan$^1$, Ken Goldberg$^{1,2}$% <-this % stops a space
\thanks{$^1$ Department of Electrical Engineering and Computer Sciences; {\small \{mdlaskey,iamwesleyhsieh, ftpokorny, anca\}@berkeley.edu}}%
\thanks{$^2$ Department of Industrial Engineering and Operations Research; {\small goldberg@berkeley.edu}}%
\thanks{$^{1-2}$ University of California, Berkeley;  Berkeley, CA 94720, USA}%
}
\begin{document}

@@

In model-free robot learning from demonstration, a robot learns to perform a task, such as driving or grasping an object in a cluttered environment as shown in Fig. \ref{fig:teaser}, from examples provided by a competent supervisor, usually a human. The problem is model-free if  the system dynamics are unknown, eg. dynamics model may be highly non-linear or noisy. In learning from demonstration, the cost function, which is the negative of the reward function, is usually unknown or significantly delayed. In the grasping for clutter example, shown in Fig. \ref{fig:teaser}(b) the robot is only penalized when an object is knocked off the table, which means a potentially large number of states can be visited before any feedback. To potentially avoid this a supervisor provides guidance through demonstrations and the robot learns a policy mapping states to controls~\cite{argall2009survey}.

@@

In SHIV, we use a query selection method that allows for a non-stationary distribution, evaluating risk based on novelty (is the state an outlier to the training distribution), as in Fig. \ref{fig:support_example}(c). Additionally, we also assign higher risk to  historically mislabeled regions, as in Fig. \ref{fig:support_example}(d). We then use a classification method based on the One Class SVM \cite{scholkopf2001estimating} to predict if a state is risky or not risky, which we we define as safe.%Our approach builds on the One Class SVM method for approximating quantile level sets \cite{}, which has already been used in novelty detection on high-dimensional image data \cite{}.

@@

We assume a known state space and set of controls. We assume access to a robot or simulator, such that we can start in some initial state and apply a series of controls and observe the following states that occur.  Lastly, we assume access to a supervisor that can provide a series of demonstrated trajectories and can be given a state visited by the robot or simulator and provide a control signal of what the supervisor would do.

@@

Providing correct control inputs or ``labeling'' at each
iteration and each state in DAgger can be tedious, imposing a large burden on the supervisor.
Instead of asking the supervisor for labels at all visited states $\{\bx_0,..,\bx_T\}$, SHIV uses a measure of risk to actively decide whether a label is necessary. For every state encountered $\bx_t$, it applies a decision function $g_{\sigma}$, which we introduce in the next section, parametrized by a risk-sensitivity parameter $\sigma$ which will capture how smooth we expect the policy function to be: smoother functions imply that training examples have larger support, and data nearby will be less risky.

SHIV only asks for supervision on states for which $g_{\sigma}(\bx) \neq 0$:

@@

\vspace{-2ex}
\begin{align}\label{eq:decision_func}
g(\bx) = \mbox{sgn}(\langle w,\Phi(x) \rangle-\rho).
\end{align}

@@

where $\alpha_i$ corresponds to the dual variables.    For training data in $\mathcal{X}$, we consider the quantile super-levelset for a user-defined threshold $\mu$ as a natural region within which
a trained policy $\pi_{\theta}$ can likely be learned since a sufficient data-density is available. The novelty detection method can be visualized in Fig. \ref{fig:support_example}(c). 
However, even when sufficient data is available, the associated control inputs may be inconsistent or noisy and a resulting policy
optimizing Eq. \ref{eq:super_objj} may still incur a large surrogate loss. To account for this, we propose a
modification to the One Class SVM:

@@

We divide up our data in to two sets those correctly classified:
$$\mathcal{D}_{s}=\{\lbrace \bx_i,\bu_i \rbrace \in \mathcal{D}_{k}, y=1\}$$
and those states incorrectly classified: 
$$\mathcal{D}_{r}=\{\lbrace \bx_i,\bu_i \rbrace \in \mathcal{D}_{k}, y=-1\}$$
A separate One-Class SVM is then trained on each set of states, ($D_{s}$ and $D_{r}$) and providing measures of the level sets, $g_{s}$ and $g_{r}$. Specified by parameters $(\nu,\sigma)$ and $(\nu_r,\sigma_r)$, respectively.

We then define the overall decision function as:
$$g_{\sigma}(\bx) = g_{r} (\bx)+ g_{s}(\bx)$$
points are deemed risky if $g_{\sigma}(\bx) \neq 0$.  Practically, this modification corresponds to
`carving out holes' in the estimated quantile super-levelset such that neighborhoods around states with $y_i=-1$ are
excluded from the super-levelset. An illustration of this can be seen in Fig. \ref{fig:support_example}(d).

@@

Our modified One Class SVM contains two different $\nu$ parameters, $\nu$ and $\nu_r$. We set $\nu = 10^{-3}$ and $\nu_r = 0.1$ for all experiments.  $\nu = 0.1$ corresponds to looking at most $90\%$ of the data, which can make it robust to outliers. We tuned $\sigma$ and $\sigma_r$ by performing a grid search over different values on the surrogate loss for a single trial of SHIV for 3 iterations for each example. Since the Driving Simulator worked with a much larger feature space the grid search started at 200, the Grasping and Surgical experiments started at 5.

@@

The supervisor drives around the track twice. We collect raw images of the simulation from a 2D bird's eye view
and use Gaussian Pyramids to down-sample the images to $125 \times 125$ RGB pixels and then extract Histogram of
Oriented Gradients (HOG) features using OpenCV. This results in a $27926$ dimensional state space description.
For both DAgger and \acro, we use a Linear Support Vector Machine (SVM) to parameterize allowable
policies $\pi_{\theta}$, with $\gamma=0.01$ as a regularization term on the slack variables, which was set via cross
validation on the initial training examples. For the optimization of Eq. \ref{eq:dual_sup} in \acro, we used an
exponential kernel with a bandwidth of $\sigma = 200$ and $\sigma_r = 200$.

@@

\textbf{Comparison to active learning approaches:}
To compare our method against a traditional confidence based active learning approach, we use a confidence measure  of distance from the hyperplane \cite{tong2002support}. The threshold was set to the average distance from the hyperplane for the mis-classified points in $\mathcal{D}_0$, which consisted of two demonstrations from our solver. We further compared against Query By Committee, which was trained via bagging \cite{breiman1996bagging}. Our data set was uniformly divided up into 3 overlapping subsets each with $80\%$ of the data then for each subset a Linear SVM was trained on. If the three classifiers agreed with each other the point was determined low risk and if they disagree it was determined high risk. We finally compared our combined risk approach against the traditional One Class SVM without the "carving out". We measured the performance in terms of the percentage of risky states that  are estimated to be safe by the active learning technique roll out. Results  are averaged over 50 tracks and shown in Fig. \ref{fig:active_comp}.

@@

\begin{figure}[t]
\centering
\includegraphics[width=9cm, height=4cm]{figures/risk_bar.eps}
\caption {  A comparison of different active learning approaches in terms of the percentage of risky states that are estimated to be safe by the active learning technique during the first policy roll out. We compare against a confidence estimate of distance to hyperplane, query by committee for 3 hypothesis classifiers, the One Class SVM, and our modified One Class SVM.  Results  are averaged over 50 tracks and the policy $\pi_\theta$ is represented as a Linear SVM. 
   }

@@

\begin{figure}[t!]
\centering
\includegraphics[width=8cm, height=6cm]{figures/dagger_shiv_one_class.eps}
\caption{We compare performance for the Driving Simulator in terms of minimization of the underlying cost function $c(\bx,\bu)$, which is the  number of times the car left the track versus the number of queries made to the supervisor. In Fig. \ref{fig:car_cost}, we plot the performance of DAgger and SHIV.  Initial results, which are run for 6 iterations each and are averaged over 40 levels, shown in Fig. \ref{fig:car_cost} suggest an $70\%$ reduction in the number of queries needed for SHIV compared to DAgger,}
\vspace*{-10pt}
\label{fig:car_cost}
\end{figure}

@@

A state $\bx\in \mathcal{X}$ consisted of the 3 dimensional pose of the six objects on the table (translation and rotation), the 3 joint angles of the arm and a scalar value in the range $[0,1]$ that measured the position of the gripper, $1$ being fully closed and $0$ being opened. For our representation of $\pi_{\theta}$, we used kernelized ridge regression with the radial basis function as the kernel with the default Sci-Kit learn parameters. We defined the cost functions, $C(\bx,\bu)$, as the sum of the number of objects knocked off the table plus $10$ times the binary value indicating if the object is grasped or not. The order of magnitude difference in cost for grasping the object is to place emphasize on that portion of the task. The bandwidth parameters for SHIV were set to $\sigma_r = 5$ and $\sigma = 6$. For the $\epsilon$ term in the our risk method, we used the median in regression error which was the L2 distance between the predicted control and the true supervisor's control.

@@

\begin{figure}[t!]
\centering
\includegraphics[width=\columnwidth, height=6cm]{figures/grasp_clutter.eps}
\caption{We compare performance for Grasping in Clutter in terms of minimization of the underlying cost function $c(\bx,\bu)$, which is the sum of the number of objects knocked off the table plus 10 times the binary value indicating if the object is grasped or not. Initial results, which are averaged over 8 different trials suggest a $65\%$ reduction in the number of queries needed for SHIV compared to DAgger,}
\vspace*{-10pt}
\label{fig:grasp_cost}
\end{figure}

@@

For trials we sample a start position from a Gaussian on the translational component of $P_0$ with isotropic variance of
$0.1$ cm. The distance between initial $P_0$ and $P_G$ is roughly $3$ cm. Our cost function $c(\bx,\bu)$ is measured in
terms of Euclidean distance in translational component, which is in centimeters. We run DAgger and SHIV both for 20 iterations and average over 40 different starting positions.
Results shown in Fig. \ref{fig:needle} suggest a reduction $67\%$ in the amount of supervisor queries needed.
